---
title: "Temperature Logger Literature Review Summary"
author: "Jess S. Glanz, Joint Research Specialist"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
mainfont: SourceSansPro
header-includes:
- \usepackage[default]{sourcesanspro}
- \usepackage[T1]{fontenc}
---

```{r LoadPackages, results="hide",message=FALSE,echo=FALSE}
library(readxl)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(ggrepel)
library(png)
library(ggstatsplot)
library(egg)
library(leaflet.minicharts)
library(leaflet)
library(gt)
library(plotly)
library(ggpubr)
```

```{r LoadData1, results="hide",message=FALSE,warning=FALSE,echo=FALSE}
ls<- read.csv('ls.csv')
lst<- read.csv('LitSum_Table.csv')
lsst<-read.csv('ls_st.csv')
```

```{r Tidy Data, results="hide",message=FALSE,warning=FALSE,echo=FALSE}
ls$dummy<-1 #add dummy variable

ls<- ls %>% mutate(bb =
                      case_when(Brand == "Onset Computer Corporation" ~ "Onset",
                                Brand == "Sea-Bird Scientific" ~ "Sea-Bird",
                                Brand == "Not mentioned" ~ "Not Mentioned",
                                Brand == "RBR" ~ "RBR",
                                Brand == "ReefNet" ~ "ReefNet",
                                Brand == "Aanderaa" ~ "Other",
                                Brand == "Alec Electronics Co" ~ "Other",
                                Brand == "Aquatec" ~ "Other",
                                Brand == "Dataflow Systems Ltd" ~ "Other",
                                Brand == "DTS" ~ "Other", 
                                Brand == "Eureka Water Probes" ~ "Other",
                                Brand == "Gemini Data Loggers Ltd" ~ "Other",
                                Brand == "General Oceanics" ~ "Other",
                                Brand == "Hugrun (now Star-Oddi)" ~ "Other",
                                Brand == "Innovasea VR2Tx " ~ "Other",
                                Brand == "Lowell Instruments" ~ "Other",
                                Brand == "Marine Geophysics Lab" ~ "Other",
                                Brand == "Maxim Integrated Products" ~ "Other",
                                Brand == "Nortek" ~ "Other",
                                Brand == "Opuhala " ~ "Other",
                                Brand == "Ryan Industries" ~ "Other",
                                Brand == "Precision Measurement Engineering Inc" ~ "Other", 
                                Brand == "Royal Netherlands Institute for Sea Research" ~ "Other",
                                Brand == "Sensornet" ~ "Other",
                                Brand == "Teledyne Marine" ~ "Other",
                                Brand == "U Beratherm and Grant Squirrel" ~ "Other",
                                Brand == "Vemco" ~ "Other",
                                Brand == "Yellow Springs Instruments" ~ "Other"),
                    .after=Brand)

```

```{r BrandFreq, results="hide",message=FALSE,warning=FALSE, echo=FALSE}
my_tab <- table(ls$Brand)                                  # Create frequency table
my_tab_prob <- (my_tab / sum(my_tab))*100

unique(ls$DOI) #Total studies: 329
length(ls$DOI) #Total studies with separation when more than one logger was used in a study: 397
#median(df$Freq1)

df <- ls %>%
  filter(duplicated(DOI) == TRUE)

unique(df$DOI) #Total studies: 51



```

## Quick Stats

```{r RenameColumns, results="show",message=FALSE,warning=FALSE, echo=FALSE}
gt_tbl <- 
  gt(lst) |>
  tab_spanner(
    label = "Shading",
    columns = c(Mentioned, Brand..Onset)
  ) |>
  tab_spanner(
    label = "Depth",
    columns = c(X.5m, Not.mentioned)
  ) |>
   tab_spanner(
    label = "Frequency",
    columns = c(Median..minutes., Mean..minutes.,Not.mentioned.1)
  ) |>
   tab_spanner(
    label = "Calibrated",
    columns = c(Yes)
  ) |>
   tab_spanner(
    label = "Data Available",
    columns = c(Yes.1)
  ) |>
  cols_label(
    X = "",
    Brand..Onset = "Brand: Onset",
    X.5m = "≤5m",
    Not.mentioned = "Not mentioned",
    Median..minutes. = "Median (minutes)",
    Mean..minutes. = "Mean (minutes)",
    Not.mentioned.1 = "Not mentioned",
    Yes.1 = "Yes"
  )

# Show the gt table
gt_tbl
```

```{r blah, results="show",message=FALSE,warning=FALSE,echo=FALSE}
lsb1 <- ls %>%
  group_by(bb) %>%
  summarise(count=sum(dummy))

lsb1$p<-lsb1$count/(sum(lsb1$count))*100
```

## What loggers are being used?

More than 60% of reviewed literature used loggers made by Onset Computer Corporation. The dominant sensor was from the HOBO Pendant series with HOBO Pro's closely following. Nearly 20% of studies did not identify which HOBO logger they used ("Not mentioned", "HOBO").

Of the studies using Sea-Bird Scientific, the second most common logger brand, models SBE 37 and 39 made up about half.

```{r blah1, results="show",message=FALSE,warning=FALSE,echo=FALSE, fig.width=8, fig.height=4}
lsb1%>%
mutate(bb = factor(bb, levels=c("Onset","Sea-Bird","RBR","ReefNet","Other","Not Mentioned"))) %>%
ggplot(aes(bb,p, fill=bb)) +  
  geom_col() +
  scale_fill_manual(values = c("navy","lightblue", "darkred","darksalmon","darkorchid","orange"))+
   theme_bw()+
  theme(legend.position="none")+
  ylab("Percentage of Total Studies")+
  xlab("Logger Brand")+
  annotate(
    geom = "text", x="Other",y=60,
    label = "*'Other' groups 24 brands that were used in less than 1% of studies", hjust = 0.50, vjust = -0.5, size = 3, color="darkorchid")
```

```{r blah2, results="show",message=FALSE,warning=FALSE,echo=FALSE}
on<-ls %>% filter(Brand=="Onset Computer Corporation")

on$model[grepl("Not",on$model)]<-"Not mentioned"

on1<- on %>% mutate(mod =
                           case_when(
                                     model=="Not mentioned"~"Other",
                                     model=="HOBO DO Data Logger"~"Other",
                                     model=="Stowaway XTI"~"Other",
                                     model=="HOBO Aquapro"~"Other",
                                     model=="Water level data logger"~"Other",
                                     model=="Optic Stowaway"~"Other",
                                     model=="HOBO Stowaway Tidbit"~"Other",
                                     model=="HOBO TidbiT"~"Tidbit",
                                     model=="HOBO Pro"~"Pro",
                                     model=="HOBO Pendant"~"Pend",
                                     model=="HOBO MXTemp"~"Other",
                                     model=="HOBO MXTemp/Light"~"Other",
                                     model=="HOBO Pendant Temp"~"Pend",
                                     model=="HOBO Pendant Temp/Light"~"Pend-LT"))


onm <- on1 %>%
  group_by(mod) %>%
  summarise(count=sum(dummy))

onm$p<-onm$count/(sum(onm$count))*100

#Compute the cumulative percentages (top of each rectangle)
onm$ymax = cumsum(onm$p)

#Compute the bottom of each rectangle
onm$ymin = c(0, head(onm$ymax, n=-1))

#Compute label position
onm$labelPosition <- (onm$ymax + onm$ymin) / 2
#
#Compute a good label

onm$label <- paste0(onm$mod, "\n  ",onm$count)#"\n  " put space between the plastic type and its count

onc<-c( "#2171B5", "#08519C", "#08306B","black","#4292C6")



###Seabird
sb<-ls %>% filter(Brand=="Sea-Bird Scientific")


sbm <- sb %>%
  group_by(model) %>%
  summarise(count=sum(dummy))

sbm$p<-sbm$count/(sum(sbm$count))*100

#Compute the cumulative percentages (top of each rectangle)
sbm$ymax = cumsum(sbm$p)

#Compute the bottom of each rectangle
sbm$ymin = c(0, head(sbm$ymax, n=-1))

#Compute label position
sbm$labelPosition <- (sbm$ymax + sbm$ymin) / 2
#
#Compute a good label

sbm$label <- paste0(sbm$model, "\n  ",sbm$count)#"\n  " put space between the plastic type and its count


sbc<-c( "#C7E9B4", "#7FCDBB", "#41B6C4", "#1D91C0", "#225EA8", "#253494" ,"#081D58","#BDBDBD", "#969696", "#737373" ,"#525252" ,"#252525", "#000000")

```

```{r blah3, results="show",message=FALSE,warning=FALSE,echo=FALSE,fig.width=4, fig.height=4}
# Make the plot
j<-ggplot(onm, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=mod)) +
  geom_rect() +
  geom_text( x=4.5, aes(y=labelPosition, label=label,color=mod), size=2) +
  scale_fill_manual(values=onc)+
  scale_color_manual(values=onc)+
  coord_polar(theta="y") + #turns bar plot into a ring shape
  xlim(c(1, 4)) +
 # theme(legend.position="none") +
  guides(fill = "none",color="none")+
  theme_void() #gets rid of gray background in plot

#Add company logo to center of donut plot
logo <- img<-readPNG("onset.png")
j

grid::grid.raster(logo, x = 0.5, y = 0.5, just = c('center','center'), width = unit(1.5, 'inches'))#x and y are values between 0 and 1 with 0 being the far bottom or left and 1 being the far top or right of the plot, width controls the size of the logo

# Make the plot
k<-ggplot(sbm, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=model)) +
  geom_rect() +
  geom_text( x=4.5, aes(y=labelPosition, label=label,color=model), size=2) +
  scale_fill_manual(values=sbc)+
  scale_color_manual(values=sbc)+
  coord_polar(theta="y") + #turns bar plot into a ring shape
  xlim(c(1, 4)) +
  # theme(legend.position="none") +
  guides(fill = "none",color="none")+
  theme_void() #gets rid of gray background in plot

#Add company logo to center of donut plot
logo <- img<-readPNG("Seabird.png")
k

grid::grid.raster(logo, x = 0.5, y = 0.5, just = c('center','center'), width = unit(2, 'inches'))#x and y are values between 0 and 1 with 0 being the far bottom or left and 1 being the far top or right of the plot, width controls the size of the logo
```

## How are they being used?

### For what and how often are measurements logged?

Studies were classified under 5 types, most of which were Ecology. Loggers were set to record measurements as low as once per 0.5 seconds (Physical Oceanography) to once per day (Biogeochemistry, Ecology), although ecological studies use a greater range of logging frequencies than any other study type.

```{r blah4, results="show",message=FALSE,warning=FALSE,echo=FALSE}

lsf<-lsst%>%filter(Freq1!="Not mentioned")
lsf$Freq1<-as.numeric(lsf$Freq1)


# sample size
sample_size = lsf %>% group_by(Study.Type) %>% summarize(num=n())

blah = lsf %>% group_by(Study.Type,Freq1) %>% summarize(n=n())

blah1 = lsst %>% group_by(Study.Type,Freq1) %>% summarize(n=n())

blah1<-blah1 %>% filter(Freq1=="Not mentioned")
b1w<-spread(blah1,Freq1,n)
```

```{r blah5, echo=FALSE, fig.height=5, fig.width=7.5, message=FALSE, warning=FALSE, results="show"}

#blah %>%
#  left_join(sample_size) %>%
#  mutate(myaxis = paste0(st, "\n", "n=", num)) %>%
 # ggplot( aes(x=Freq1, y=n)) +
#  geom_area(aes(fill=Study.Type), alpha=0.4) +
  #geom_line(aes(color=Study.Type),linewidth=0.5) +
 # 
 # scale_color_manual(values=c("#453781FF","#287D8EFF","pink","#3CBB75FF","orange")) +
  #geom_point(aes(color=Study.Type), alpha=0.9) +
#  theme(
 #   legend.position="none",
  #  plot.title = element_text(size=11)
#  ) +
# theme_bw()+
#  guides(fill=FALSE, color=FALSE)+
 # xlab("Frequency 1 (minutes)")+
#  ylab("Count")
 # facet_wrap(~Study.Type)

#my_tag <- c("A) Biogeochemistry, n=27", "B) Ecology, n=249", "C) Physical Oceanography, n=56", "D) Management, n=3", "E) Data Validation, n=2")
#tag_facet(p, 
 #         x = 0, y = 55, 
  #        vjust = 0, hjust = 0.03,
   #       open = "", close = "",
    #      fontface = 4,
     #     size = 3,
      #    family = "serif",
       #   tag_pool = my_tag)


###
p1<-blah %>%
  left_join(sample_size) %>%
  mutate(myaxis = paste0(Study.Type, "\n", "n=", num)) %>%
  ggplot( aes(x=Freq1, y=myaxis, fill=Study.Type)) +
    geom_violin(width=1.4) +
    geom_point() +
   scale_fill_manual(values=c("#453781FF","#287D8EFF","pink","#3CBB75FF","orange")) +
    theme_classic() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
   # ggtitle("A Violin wrapping a boxplot") +
    ylab("")+
  xlab("Sampling Frequency (minutes)")

p2<-blah1%>%
ggplot(aes(fill=Study.Type, y=Freq1, x=n)) + 
    geom_bar( stat="identity")+
  scale_fill_manual(values=c("#453781FF","pink","orange")) +
  # scale_y_discrete(limits=rev)+
  xlab("")+
  ylab("")+
  guides(fill=FALSE)+
  #scale_y_discrete(labels=myylab)+
   theme_classic()+ #gets rid of gray background in plot
   geom_text(aes( x=12, y="Not mentioned", label=paste0("Phys Oceanography", "\n", "n=", 23)),   
           color="white", 
           size=3 , angle=15 )+
     geom_text(aes( x=60, y="Not mentioned", label=paste0("Ecology", "\n", "n=", 73)),   
           color="white", 
           size=3 , angle=15 )+
   geom_text(aes( x=100, y="Not mentioned", label=paste0("Biogeochem", "\n", "n=", 10)),   
           color="white", 
           size=3 , angle=15 )

ggarrange(p2,p1,ncol=1,nrow=2,align="v",heights = c(0.7, 2),labels = c("A", "B"))




```

### Geographic distribution of studies, their logger brands, and which studies mentioned shading or protecting loggers

This map is interactive. Zoom in and out, click and drag to move around, and click on a pie chart or icon to learn more. The pie chart info is also shown in the figure below the map with regions descending from most to least number of studies.

```{r blah6, results="show",message=FALSE,warning=FALSE,echo=FALSE}
lsbs <- ls %>%
  group_by(country,f.lat,f.lon,bb) %>%
  summarise(count=sum(dummy))

lsbsw<-spread(lsbs,bb,count)
lsbsw[is.na(lsbsw)] <- 0
lsbsw$Total<-lsbsw$`Not Mentioned`+lsbsw$Onset+lsbsw$Other+lsbsw$RBR+lsbsw$ReefNet+lsbsw$`Sea-Bird`

colors <-c("navy","lightblue", "darkred","darksalmon","darkorchid","orange")



shaded<-ls %>% filter(Shaded=="Yes")

ds <- shaded %>%
  filter(duplicated(DOI) == FALSE)

icon = makeIcon("shaded2.png", "shaded2.png", 50, 50)
```

```{r blah7, echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE, results="show"}

leaflet(data=ds) %>% 
  addTiles() %>% 
  setView( lng = 0, lat = 0, zoom = 2 ) %>% 
  addProviderTiles("Esri.WorldGrayCanvas") %>%
  addMarkers(~long, ~lat,  icon=icon, popup =  ~paste(country, "<br>Brand:", Brand,
                                                               "<br>Model:", Model,
                                                               "<br>Study Type:", Study.Type)) %>%
addMinicharts(lsbsw$f.lon, lsbsw$f.lat, layerId = lsbsw$country, chartdata = lsbsw[, c("Onset","Sea-Bird","RBR","ReefNet","Other","Not Mentioned")], 
              type = "pie",
              colorPalette = colors,
              width = 100 * sqrt(lsbsw$Total) / sqrt(max(lsbsw$Total)))

```

```{r blah8, echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE, results="hide"}

ls<- ls %>% mutate(st =
                         case_when(Study.Type == "Physical Oceanography" ~ "Physical Oceanography",
                                   Study.Type == "Physical Oceanography?" ~ "Physical Oceanography",
                                   Study.Type == "Physical Oceanography? " ~ "Physical Oceanography",
                                   Study.Type == "Management" ~ "Management",
                                   Study.Type == "Management/Data validation" ~ "Management",
                                   Study.Type == "Management" ~ "Management",
                                   Study.Type == "remote data validation" ~ "Physical Oceanography",
                                   Study.Type == "Ecology" ~ "Ecology",
                                   Study.Type == "Ecology/Physical Oceanography?" ~ "Ecology/Physical Oceanography",
                                   Study.Type == "Biogeochemistry" ~ "Biogeochemistry",
                                   Study.Type == "Physical Oceanography/Biogeochemistry" ~ "Physical Oceanography",
                                   Study.Type == "Ecology/Biogeochemistry" ~ "Ecology"
                                   
                         ),
                       .after=Study.Type)

sta<-ls %>% 
    group_by(bb, st) %>%
    summarize(Count = n()) %>%
    pivot_wider(id_cols = bb, names_from = st, values_from = Count,
                values_fill = 0) 

stal<-gather(sta,st,count, 2:6)
```

```{r blah11, echo=FALSE, fig.height=5, fig.width=7.5, message=FALSE, warning=FALSE, results="hide"}

#stal %>%
#  mutate(bb = factor(bb, levels=c("Onset","Sea-Bird","RBR","ReefNet","Other","Not Mentioned"))) %>%
 # ggplot( aes(x=bb, y=count,fill=bb)) +
 # geom_area(aes(fill=st), alpha=0.4) +
#  geom_line(aes(color=st),linewidth=0.5) +
  #scale_fill_manual(values=colors) +
 # scale_color_manual(values=c("#453781FF","#287D8EFF","pink","#3CBB75FF","orange")) +
# geom_bar(position="dodge", stat="identity") +
 # theme(
  #  legend.position="none",
   # plot.title = element_text(size=11),
    #axis.text.x = element_text(angle=45, vjust=.5, hjust=1)
  #) +
 #theme_bw()+
#  guides(fill=FALSE, color=FALSE)+
 # xlab("")+
  #ylab("Count")+
  #facet_wrap(~st)

```

```{r blah9, echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE, results="hide"}
cb<-ls %>% 
    group_by(country, bb) %>%
    summarize(Count = n()) %>%
    pivot_wider(id_cols = country, names_from = bb, values_from = Count,
                values_fill = 0) 

cb$Total<-cb$`Not Mentioned`+cb$Onset+cb$Other+cb$RBR+cb$ReefNet+cb$`Sea-Bird`

cbl<-gather(cb,Brand,Count,2:7)
cbl$p<-(cbl$Count/cbl$Total)*100
cbl2<-cbl%>%select(1:3,5)

cbw<-spread(cbl2,Brand,p)
cbw2<-cbw %>%
    group_by(country) %>%
    summarise_all(na.omit)
cbw2<-cbw2 %>% arrange(desc(Total))

```

```{r blah10, echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE, results="hide"}
colors <-c("navy","lightblue", "darkred","darksalmon","darkorchid","orange")

myylab <- paste(as.factor(cbl2$country), "\n(N=", cbl2$Total, ")", sep="")

```

```{r blah12, echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE, results="show"}
p<-cbl2%>%
mutate(country = factor(country, levels=c("Australia","Florida","Hawaii","French Polynesia","Caribbean","Mexico","Japan","Line Islands","Saudi Arabia",
         "Panama","China","United Arab Emirates","Palau","Thailand","Guam","Samoa","Texas","Bermuda","Brazil","Kenya","Taiwan","Chagos Archipelago","Jordan","New Caledonia","Philippines","Singapore","Belize","Papua New Guinea","South Africa","Bahamas","Ecuador","Egypt","Maldives","Vietnam","Cape Verde","Chile","Colombia","Costa Rica","Curacao","Fiji","India","Indonesia","Israel","Madagascar","Mauritius","Micronesia","Qatar","South Korea","Tobago","Vanuatu","Zanzibar"))) %>%
ggplot(aes(fill=Brand, y=country, x=p)) + 
    geom_bar(position="fill", stat="identity")+
  scale_fill_manual(values=colors,breaks=c("Onset","Sea-Bird","RBR","ReefNet","Other","Not Mentioned"))+
   scale_y_discrete(limits=rev)+
  xlab("")+
  ylab("")+
  guides(fill=FALSE)+
  #scale_y_discrete(labels=myylab)+
   theme_classic() #gets rid of gray background in plot

ggplotly(p)#make plot interactive


```
